---
title: "ANOVA example 1"
author: "Petri Haavisto"
format: html
editor: visual
---

## ANOVA - Chicken Weights by Feed Type

Analysis of variance, or ANOVA, is used to test whether a quantitative variable and categorical variable are independent. For instance, doctors may wonder whether people in different age groups respond equally to a new blood pressure medicine, or whether several different medications are all equally effective.

We are using chicken weights \~ feed as an example.

Libraries etc...

```{r}
#| message: false
library(conflicted)  
library(tidyverse)
#library(broom)
#library(GGally)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
theme_set(theme_minimal()) #set preferred theme for plots
```

Chicken are given different types of fee. We are interested, how this is affecting their weight. Is weight independent from feed? Or is it affected by feed?

```{r}
view(chickwts)
```

### Visualize data

Always good to start examine data visually.

```{r}
ggplot(chickwts , aes(x = feed, y = weight))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter()
```

One good thing about visualization is that we can roughly see, if the assumptions have been met:

1.  Variances between groups should be similar in other words width of the boxes should be roughly same.

2.  Data should be normally distributed between groups. So these boxes should be relatively symmetric.

3.  Observations should be independent from each other. This is something you can't check with plot above.

### Run the analysis

```{r}
model <- aov(chickwts$weight ~ chickwts$feed)
summary(model)
```

What we care here the most is the p-value. This is the probability that **we got F value just by random change** if null hypothesis were true. We have null hypothesis that group means are all same in the larger population. Small p-value is strong evidence against the p-value.

ANOVA does not give us information, how big the differences are between groups. In order to find out differences, we have to run a post-hoc analysis ("after the event"). One widely used test is the TukeyHSD.

### Post-hoc test: Tukey Honest Significant differences

This runs significance test pairwise and creates set of confidence intervals on the differences.

```{r}
TukeyHSD(model)
```

Here we can see that differences between horsebean and casein are big. On the other hand differences between meatmeal and casein are not.
