---
title: "Confidence Intervals"
author: "Petri Haavisto"
format: html
editor: visual
---

## Confidence intervals - CO2

We are going to use data set CO2 for this. Let's first run t-test.

```{r}
view(CO2)
t.test(CO2$uptake)
```

We are interested in 95 percent confidence intervals that this test gives us. Test does not give us **margin of error** but we can compute that. margin of error = upper limit - sample mean

```{r}
upper_lim <- 29.55997
samp_mean <- 27.2131
marg_of_err <- upper_lim - samp_mean
marg_of_err
```

We can also calculate the margin of error with:

```{r}
qt(.975, 83)*sd(CO2$uptake)/sqrt(84)
```

What if we want different confidence intervals?

```{r}
t.test(CO2$uptake, conf.level = .90)
```

Notice that when you decrease the level of confidence you should get more narrow confidence interval.

### Calculate confidence interval manually

We will need:

-   Sample mean

-   Sample variance

-   Sample size = 84

-   Critical value

-   Confidence level (95 % in this case)

```{r}
s_mean <- mean(CO2$uptake) 
#s_vari <- var(CO2$uptake)
s_vari <- variance(CO2$uptake)
s_stdv_1 <- sqrt(s_vari)
s_stdv_2 <- sd(CO2$uptake)
#s_size <- nrow(CO2)
s_criv <- qt(.975, 83)
marg_of_error <- s_criv * (s_stdv_1/sqrt(84))
lover_limit = s_mean - marg_of_error
upper_limit = s_mean + marg_of_error
s_mean
s_vari
s_stdv_1
s_stdv_2
marg_of_error
lover_limit
upper_limit
```

Function for manual variance calculation:

```{r}
variance <- function(x){
  x = as.numeric(x)
  x = na.omit(x)
  m = mean(x)
  return(
    sum((x-m)^2, na.rm = TRUE)/(length(x) - 1)
  )
}
```

Let's test function.

```{r}
variance(CO2$uptake)
```
