---
title: "Logistic regression"
author: "Petri Haavisto"
format: html
editor: visual
---

## Logistic regression - iris

Libraries:

```{r}
#| echo: false
#| message: false
library(conflicted)  
library(tidyverse)
library(tidymodels)
library(tinytex)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
theme_set(theme_minimal()) #set preferred theme for plots
```

Prep data. Here we split data so they are either **setosas** or **not setosas**

```{r}
iris <- iris |> 
  mutate(setosa = as.integer(Species == "setosa"))
View(iris)
```

We split data to training and testing parts:

```{r}
set.seed(2)
split <- initial_split(iris, prop = .80, strata = setosa)
iris_train <- training(split)
iris_test <- testing(split)
#View(iris_test)
```

Visualize the data and check for assumptions:

```{r}
ggplot(iris_train, aes(x = Sepal.Length,
                       y = setosa))+
  geom_jitter(height = .05, alpha = .5)
```

Looks like this data is goo candidate for logistic regression model. Same with regression line:

```{r}
#| message: false
ggplot(iris_train, aes(x = Sepal.Length,
                       y = setosa))+
  geom_jitter(height = .05, alpha = .5)+
  geom_smooth(method = "glm", method.args =  list(family = "binomial"),
              se = FALSE)
```

A Logistic regression model for binary response variable *y* with a quantitative explainer *x* has the following form:

$$
\begin{align*}
\begin{cases}
  logit(p) = \beta_0 + \beta_1x \quad &\textit{(systematic component)}\\
  y \sim bin(p, 1) \qquad &\textit{(random component)}
  \end{cases}
\end{align*}
$$

Such a model gives the probability that a new observation will be success given its *x*-value. R fits the coefficients $\beta_0$ and $\beta_1$ using maximum likelihood estimation. Here $p$ is probability to being success.

### Building the model

```{r}
model <- glm(setosa ~ Sepal.Length,
             data = iris_train,
             family = "binomial")
summary(model)
```

Above intersection is in logit scale (29.210) so it is not intersection on the same scale that plot had. Similar way -5.413 is not slope on the plot but slope on the logit scale. So what we have been given here are the $\beta_0$ and $\beta_1$ in that model form. So the model given by R is:

$$
\begin{align*}
logit(p) &= 29.2 - 5.4 \times Sepal.Length\\\\
ln \left(\frac{p}{1-p}\right) &= 29.2 -5.4 \times Sepal.Length\\\\
p &= \frac{e^{29.2-5.4\times Sepal.Length}}{1 + e^{29.2-5.4\times Sepal.Length}} 
\end{align*}
$$

Now we can try values for this formula. For instance and iris with a sepal length of 5 mm is predicted to be a setosa with approximately 90% probability.

$$
p = \frac{e^{29.2-5.4\times 5}}{1 + e^{29.2-5.4 \times 5}} 
$$

Let's make a function that can do this calculation for us:

```{r}
setosa_prob <- function(len){
   top <- exp(1)^(29.2-5.4*len)
   bottom <- 1  + exp(1)^(29.2-5.4*len)
   return(top/bottom)
}
```

Call:

```{r}
result <- setosa_prob(5)
result
```

If you compare this value to the plot, we can see that probability at x = 5 is indeed about 0.90.

### Evaluate the model on the testing set

```{r}
iris_test <- iris_test |> 
  mutate(setosa_prob = predict(model, iris_test, type = "response"),
         setosa_pred = ifelse(setosa_prob > .5, 1, 0))
View(iris_test)
```
