---
title: "Beautiful tables with gt"
author: "Petri Haavisto"
format: html
editor: visual
---

## Beautiful tables with the gt package

Libraries:

```{r}
#| message: false
library(conflicted)  
library(tidyverse)
library(gt)
library(modeldata)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```

We are using data set "scat".

```{r}
View(scat)
```

First little bit processing of the the data. This creates brilliant recap of variables grouped by species.

```{r}
scat_table1 <- scat |> 
  select(Species,
         Site,
         Length,
         Diameter,
         Mass,
         d13C,
         d15N) |> 
  group_by(Species) |> 
  summarize(across(where(is.numeric),
                   \(x) round(mean(x, na.rm = TRUE), 1)),
            count = n())

```

Second tibble

```{r}
scat_table2 <- scat |> 
  select(Species,
         Location,
         Site,
         Length,
         Diameter,
         Mass,
         d13C,
         d15N) |> 
  group_by(Species, Site) |> 
  summarize(across(where(is.numeric),
                   \(x) round(mean(x, na.rm = TRUE), 1)),
            Count = n())
```

### Table with gt

```{r}
scat_table1 |> 
  gt()
```

Adding/changing things manually to a table (start with "tab" to see what options is available).

```{r}
#?scat
scat_table1 |> 
  gt(rowname_col = "Species") |> 
  cols_label(count ~ "Count") |> 
  tab_header(title = "California poopie",
             subtitle = "Morphometric data on scat") |> 
  tab_source_note(md("**Source**: Reid, R. E. B. (2015). A morphometric modeling approach to distinguishing among bobcat, coyote and gray fox scats. *Wildlife Biology*, 21(5), 254-262")) |> 
  tab_spanner(label = "Physical averages", columns = Length:Mass) |> 
  tab_stubhead(label = md("**Species**"))
```

Notice that above we used "md" markdown system. You can also use html system in gt-tables.

Also check out: ?footnote and ?fmt... (different formatting things).

### Stylized version

```{r}
scat_table1 |> 
  gt() |> 
  cols_label(count ~ "Count") |> 
  tab_header(title = "California poopie",
             subtitle = "Morphometric data on scat") |> 
  tab_source_note(md("**Source**: Reid, R. E. B. (2015). A morphometric modeling approach to distinguishing among bobcat, coyote and gray fox scats. *Wildlife Biology*, 21(5), 254-262")) |> 
  tab_spanner(label = "Physical averages", columns = Length:Mass) |> 
  opt_stylize(style = 2, color = "blue")
```

### Grouped data

This is going to work when we use "ungroup" command

```{r}
scat_table2 |> 
  ungroup() |> 
  gt() |> 
  opt_stylize(style = 3)
```
